---
title: "GSE161533 Analysis"
output: html_notebook
---

```{r}
if (!requireNamespace("BiocManager", quietly = TRUE))
    install.packages("BiocManager")

BiocManager::install(c("affy", "affyPLM", "limma", "gage", "gageData"))

library(affy)
library(affyPLM)
library(limma)
library(gage)
library(gageData)

# Read the .CEL file
cel_data <- ReadAffy(filenames="GSM4909596_20200609-190360C_121P_HG-U133_Plus_2_.CEL")

```
```{r}
head(cel_data)
```

```{r}

# Perform RMA normalization
normalized_data <- rma(cel_data)

# Get expression values
expression_matrix <- exprs(normalized_data)

# Basic quality control plots
par(mfrow=c(2,2))
hist(expression_matrix, main="Distribution of Normalized Expression Values", xlab="Expression")
boxplot(expression_matrix, main="Boxplot of Normalized Expression Values", ylab="Expression")
plot(density(expression_matrix), main="Density Plot of Normalized Expression Values")
qqnorm(as.vector(expression_matrix))
qqline(as.vector(expression_matrix), col="red")

# Load KEGG pathways
data(kegg.gs)

# Perform pathway analysis
pathway_results <- gage(expression_matrix, gsets=kegg.gs, ref=NULL, samp=NULL)

# Get the top up-regulated and down-regulated pathways
top_pathways <- rbind(head(pathway_results$greater, 10), head(pathway_results$less, 10))

# Print top pathways
print(top_pathways)

# Plot top pathways
par(mar=c(5,20,4,2))
barplot(top_pathways$stat, 
        names.arg=rownames(top_pathways), 
        horiz=TRUE, 
        las=1, 
        main="Top 10 Up and Down Regulated Pathways",
        xlab="t-statistic")

# Heatmap of top genes in a specific pathway (e.g., "KEGG_CELL_CYCLE")
cell_cycle_genes <- kegg.gs[["KEGG_CELL_CYCLE"]]
cell_cycle_expr <- expression_matrix[cell_cycle_genes,]
heatmap(cell_cycle_expr, main="Expression of Cell Cycle Genes")

# Save results
write.csv(pathway_results$greater, "upregulated_pathways.csv")
write.csv(pathway_results$less, "downregulated_pathways.csv")

```

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.
